---
title: "Final_Project"
author: "Cameron Chambers, Bram Stults, Priscilla Wen"
date: "2024-04-02"
output: html_document
header-includes:
   - \usepackage{multirow}
editor_options: 
  chunk_output_type: console
urlcolor: blue
---

# Introduction
//Begin paper here

### Background 
//Discuss background of our project here


# Mathematics behind Divisive Hierarchical Clustering
//Mathematics information here


# Analysis on our Dataset
//Analysis here


# Diagnostic and Remedial Measures
//Diagnostics here


# Strengths, Weaknesses, and Uses
//Information regarding our ML model here


# Conclusion
//Conclusion of paper here


# References
//References used here


# RShiny App
``` {r}
library(dendextend)
library(ggplot2)
library(plotly)
library(shiny)
library(factoextra)
library(cluster)
library(stats)

data = read.csv("Studio Ghibli [TEST].csv", header = TRUE, row.names = 1)

data_new <- na.omit(data)
data_new <- data_new[c("Year", "Budget", "Revenue", "Duration")]
col_names <- c("Year", "Budget_in_$M", "Revenue_in_$", "Duration_in_Minutes")
data_new <- setNames(data_new, col_names)
data_new <- matrix(data_new)

data_new

clean_data <- apply(data_new, 2, function(column) {
  if(any(is.na(column))) {
    column[is.na(column)] <- mean(column, na.rm = TRUE)
  }
  return(column)
})

dend <- data_new %>%
  dist() %>%
  hclust(method = "ave") %>%
  as.dendrogram()
dend2 <- color_branches(dend, 5)

p <- ggplot(dend2, horiz = T, offset_labels = -3)
ggplotly(p)



data_new
write.csv(data_new, "Studio Ghiblie [TEST2].csv")
```

``` {r}
dist_matrix <- dist(data_new)
hc <- hclust(dist_matrix)
dend <- as.dendrogram(hc)
dend_colored <- color_branches(dend, k = 4)  
plot(dend_colored, main = "Colored Dendrogram")
legend("topright", legend = unique(dend_colored$labels$color_groups), 
       fill = unique(dend_colored$labels$color_groups), title = "Cluster")

# If you prefer ggplot2 for better customization:
ggplot(dend_colored, horiz = TRUE) + theme_minimal() +
   geom_segment(aes(x = x, y = y, color = color_groups)) +
   scale_color_manual(values = unique(dend_colored$labels$color_groups)) +
   theme(legend.position="top")
```

``` {r}
# Load necessary libraries
library(shiny)
library(shinydashboard)
library(stats)

# Define UI for the application
ui <- dashboardPage(
  dashboardHeader(title = "Dendrogram Viewer"),
  dashboardSidebar(
    fileInput("dataFile", "Choose CSV File",
              accept = c(
                "text/csv",
                "text/comma-separated-values,text/plain",
                ".csv")),
    tags$hr(),
    checkboxInput("header", "Header", TRUE),
    radioButtons("sep", "Separator",
                 choices = c(Comma = ",", Semicolon = ";", Tab = "\t"),
                 selected = ",")
  ),
  dashboardBody(
    plotOutput("dendrogramPlot"),
    tags$hr(),
    verbatimTextOutput("fileInfo")
  )
)

# Define server logic
server <- function(input, output) {
  data <- reactive({
    inFile <- input$dataFile
    if (is.null(inFile)) {
      return(NULL)
    }
    df <- read.csv(inFile$datapath, header = input$header, sep = input$sep, na.strings = c("NA", "na", ""))
    df
  })

  output$fileInfo <- renderPrint({
    df <- data()
    if (is.null(df)) {
      "No file uploaded."
    } else {
      summary(df)
    }
  })

  output$dendrogramPlot <- renderPlot({
    df <- data()
    if (is.null(df)) {
      return()
    }

    clean_data <- apply(df, 2, function(column) {
      if (any(is.na(column))) {
        column[is.na(column)] <- mean(column, na.rm = TRUE)
      }
      return(column)
    })

    dist_matrix <- tryCatch({
      dist(clean_data)
    }, warning = function(w) {
      cat("Warning: ", w$message, "\n")
      return(NULL)
    }, error = function(e) {
      cat("Error: ", e$message, "\n")
      return(NULL)
    })
    
    # Execute code that should always run (similar to 'finally')
    cat("Distance computation attempted.\n")

    dend <- data_new %>%
      dist() %>%
      hclust(method = "ave") %>%
      as.dendrogram()
    dend2 <- color_branches(dend, 5)
    p <- ggplot(dend2, horiz = T, offset_labels = -3)
    ggplotly(p)
  })
}

# Run the application 
shinyApp(ui, server)
```
